---
title: "Analysis Report"
author: "MG"
date: "Modified: `r Sys.Date()`"
output: 
# bookdown::pdf_book: default
bookdown::html_document2: default
bibliography: ["ref.bib"]
csl: ieee.csl
natbiboptions: round
link-citations: TRUE
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
# for not outputting echo
knitr::opts_chunk$set(echo = FALSE, results='hide', message=FALSE, warning=FALSE)

# --------------------------------
# libraries
# --------------------------------
library(readr)
library(dplyr)
library(tidyverse)


# --------------------------------
# constants
# --------------------------------
# original resorts
PATH_RESORTS_ORG="input/mg_resorts.csv"
PATH_SNOW="input/mg_snow.csv"

# processed file
PATH_RESORTS="input/mg_resorts_gs.csv"


# functions and library loads

# Compute the percentage
# @param value Value to be presented as percentage
# @param digi What precision, how many digits after decimal point
# @return rounded value with digi digits after decimal point
percentage <- function(value, digi) {
  round(value*100, digits = digi)
}
```

# Business Task Statement

The purpose of this study is to build a dashbard to help skiers find their ideal
destination for the next winter season. The dashboard should allow for 
exploration and interactivity.

The original dataset includes details on 499 ski resorts around the world 
(location, slopes, lifts, prices, and ski season). There is also 
a table with the snow cover around the world for each month of 2022 by
latitude and longitude. It contains two data tables in CSV format and
description of fields.

```{r}
# read the csv file
dforg <- read_csv(PATH_RESORTS_ORG)
df <- read_csv(PATH_RESORTS)
df_snow <- read_csv(PATH_SNOW)
```

# Deliverables

The dashboard in Shiny.
<!-- ```{r, results='markup'} -->


# Data

The original data is located at [@maven_ski_resorts]. The data is
organized in two .csv files:

-   `mg_data_dictionary.csv` - the description of fields.
-   `mg_resorts.csv` - the actual data (original data)
-   `mg_resorts_gs.csv` - the processed file
-   `mg_snow.csv` - snow cover for selected GPS locations for the entire 2022 year.

It is not specified which year the resorts data regards. However, I expect that 
it covers 2022 year. In addition, [@maven_ski_resorts] states that 
the data source is [https://ski-resort-stats.com](https://ski-resort-stats.com) and NASA Earth Observations [@ski_resorts_source]
with License *Public Domain*.

##  mg_resorts.csv and mg_resorts_gs.csv

File | Data Structure | No. of Fields | No. of Records | Date Added |
|:--:|:--------------:|:-------------:|:--------------:|:----------:|
mg_resorts.csv |  Table |       `r ncol(dforg)`      | `r nrow(dforg)`            | 02/07/2023 |
mg_resorts_gs.csv |  Table |       `r ncol(df)`      | `r nrow(df)`            | 02/07/2023 |

After data examination, it occurred that there are duplicates (explained
later). I also reorganized fields for seasons, splitted seasons
to months, added the field 'Seasons number' indicating the number
of seasons shown in the original file. I added the numerical columns 
corresponding to last four fields with 'Yes' 'No' answers (there were
no other answers).

For analysis I use the `r PATH_RESORTS` file.


## Data Structure

The description of main fields is in `mg_data_dictionary.csv`:

For the `r PATH_RESORTS` file.

Name | Description | Notes |
|:--|:--------------|:-----------------|
| ID | Identifier for each resort | |
| Resort | Name of the ski and snowboard resort | |
| Country | Resort country location | Unique countries:  `r length(unique(df$Country))` |
| Continent | Resort continent location | Unique continents: `r length(unique(df$Continent))` |
| Price | Ski pass cost for 1 adult for 1 day in the main season in Euro | Season | Normal start and end of the ski season (weather permits) | |
| Months | 0 indicates a resort is closed, 1 the resort is open | |
| Highest point	| Highest mountain point  in meters | |
| Lowest point	| Lowest possible point to ski in meters | |
| Beginner slopes |	Total length of "children", "blue", and "green" slopes in km | |
|Intermediate slopes | Total length of "red" slopes in km | |
| Difficult slopes	| Total length of "black", "advanced", and "expert" slopes in km |
| Total slopes|	Total length of slopes in km | |
| Longest run	| Longest possible continuous run in km | |
| Snow cannons |	Total amount of snow cannons | |
| Surface lifts |	Total number of surface lifts, including T-bar, Sunkidslift, Rope lifts and people mover | |
| Chair lifts	| Total number of chair lifts | |
| Gondola lifts	| Total number of gondola lifts, including Gondola, Train lifts, Funicular, Combined gondola and chairlifts, Helicopter lifts, Snowcats and Aerial tramways | |
| Total lifts	| Total number of lifts | |
|	Lift capacity |	Number of passengers the resort's lift system can move in an hour | |
| Child friendly Org	| Is the ski resort child friendly? | Categorical |
| Child friendly | Is the ski resort child friendly? | 0 - means 'no', 1 - means 'yes' |
| Snowparks	Org| Does the resort have one or more snowparks? | Categorical |
| Snowparks	| Does the resort have one or more snowparks? | 0 - means 'no', 1 - means 'yes'  |
|	Nightskiing	Org| Does the resort offer skiing on illuminated slopes? | 0 - means 'no', 1 - means 'yes' |
|	Summer skiing	Org | Does the resort offer skiing during the summer? | Categorical |
|	Summer skiing	| Does the resort offer skiing during the summer? | 0 - means 'no', 1 - means 'yes' | |

Processed file `r PATH_RESORTS` summary
```{r, results='markup'}
summary(df)
```

<!-- If you want to make it visible despite was globally defined
```{r , results='markup'}
summary(dforg)
```
-->


There is a file `mg_snow.csv` that contains data regarding snow cover for selected GPS locations for the entire 2022 year.

Name | Description | Notes |
|:--|:--------------|-----------------|
| Month |	Date to represent the month of the year (not just the first day) | |
| Latitude	| Latitude at the center of the region (every "region" is 0.25x0.25 degrees in size) | |
| Longitude	| Longitude at the center of the region (every "region" is 0.25x0.25 degrees in size) | |
| Snow	| Percent of time the region was covered in snow during the month | |

Processed file `r PATH_SNOW` summary
```{r, results='markup'}
summary(df_snow)
```

# Data Processing


## Data Cleaning

I used mainly Google Sheets for data cleaning. The number of records is 480
and it was not an issue to process it. I used among others VLOOKUP, Data Cleanup

### Duplicates

- The below records have all fields the same values, including latitude
  and longitude. I removed 143, 145, and 146, and renamed the resort
  field in ID=140 to "4 Vallees (Verbier, Nendaz, Thyon, Veysonnaz)".

| ID | Resort |
|:--:|:-------|
| 140 |	Verbier (4 Vallees) |
| 143	| Nendaz (4 Vallees)  |
| 145	| Thyon (4 Vallees)   |
| 146	| Veysonnaz (4 Vallees) |

- Records: ID=15 and ID=134 have the same values, including country and
  longitude and latitude. Interestingly, although Cervinia is in Italy
  the country record shows Switzerland so I assumed the record ID=15 
  should remain

15	Zermatt - Matterhorn
134	Cervinia

- Records: ID=104 and 111, have the same field values apart from
  ID, Resort, and Longest Run (Anton - 10, Lech - 6, which roughly
  is what I can find on-line: 11km and 4.5km, respectively. 
  However, the total number of lifts for Lech is 34, for Anton is 38, 
  and for Warth-Schroecken is 15, 
  which gives total 87 as of Mar 17, 2023. So I assume the data 
  regards a single place Ski Arlberg that includes (Anton, Lech, and
  Warth-Schroecken). I remove ID=11, and rename St. Anton to 
  Ski Arlberg (St Anton and Lech)
  
  Also combining 120 and 123 records (Zurs, and St. Christoph)
  
|  ID |	Resort | Longest Run |
|:--:|:--:|:--:|
| 104 |	St. Anton (Ski Arlberg) | 10 |
| 111	| Lech (Ski Arlberg) | 6 |


- Records: ID=20 and 32 have many the same field values, the name is
  just reverted. The resort is in France (not in Italy as in record ID=32) and the longest run in record ID=32 is 0. I removed ID=32.

- Records: ID 19, 61, 62 seems to have many same values, so I 
  leave only one 
- Records: ID 66, 67, 69, 74, 149, 150 have very same values, only
  only 66 has the longest run > 0, so I leave it:
  
  Les 3 Vallees: Courchevel, Saint Martin de Belleville, La Tania-Val Thorens/Les Menuires/Meribel, Val Thorens, Meribel,
Les Menuires

- Records: ID 38, 72, and 133, seems values are same so inferring that the records have combined stats. Rec. 72 longest run has 0, so removing it.

- Records: 106 and 148 combined into 106 (148 removed)
- 118, 121 combined - record 121 removed.


### Resort

I used GoogleSheets to deal with special characters. It seems that it is in Windows-1252 or Latin-1 encoding. I removed question marks. 

### Country

- 38 unique values

|  Country | No. of Resorts |
|:--:|:--:|
| Austria | 84  |
| USA | 78 |
| France	| 73  |
| Switzerland | 55 |
| Italy | 42 |

### Continent

- 5 unique values

|  Continent | No. of Resorts |
|:--:|:--:|
| Europe | 341  |
| North America	| 98  |
| Asia | 24 |
| Oceania | 10 |
| South America | 7 |

### Price

The below records have price = 0.

|  ID | Resort  | Continent | 
|:--:|:--:|:--:|
| 196	| Perisher | Oceania |
| 235 | Yellowstone Club | North America |
| 335 |	Pragelato | Europe |
| 349	| Uludag-Bursa | Asia |
| 369	| Palandoken-Ejder 3200 World Ski Center- | Asia |
| 387	| High1 Resort | Asia |
| 390	| Alpika Service | Europe |
| 475	| Sun Mountain-Yabuli | Asia |
| 495	| Puigmal | Europe |


### Season

The season field can have a single range, two ranges, three ranges,  unknown or year-around value. A single range can be a single month like "April" or "December-April". Two and three month ranges does not include a single month. There are two three seasons value: "December - April, June - August, October - November", and "October - November, December - May, June - October." The second one accounts for the entire year. I transformed it with VLOOKUP and marked each month whether it is in season. If there are no gaps, 
i.e., no bordering '0' then the information about when the season starts and 
ends is lost (i.e., where there is continuity of '1'). 
I can add a field a number of seasons, though.

### Latitude and Longitude

The min latitude in [0; `r min(df$Latitude[df$Latitude >= 0])` ] = `r min(df$Latitude[df$Latitude >= 0])`  and the max
latitude in [`r max(df$Latitude[df$Latitude < 0])`, 0) = 
`r max(df$Latitude[df$Latitude < 0])`.

## Missing Values
Surprisingly there were no missing values.

## White Spaces
I checked with Google Sheets there are no white spaces.

# Data Credibility

The data is provided by a respected website devoted to Data Analytics
[@maven_ski_resorts]. 

# Data Bias

My assumption is that the data is collected and presented in a fair manner.

# License

According to [@maven_ski_resorts] the license is *Public Domain*.

# Future Work

I have not used the data related to the snow cover. Including viz and
info regarding the snow cover for resorts of interest could be valuable
for ski trip planners.

# References
